### 2023-03-13 main.tex
label
sec1

At each site $x\in V$, we describe the occupation with an $N-$dimensional vector in which the value of the $k$-th component $n_{k}^{x}$ denotes the number of particles of species $k\in \{1,\ldots,N\}$. On each vertex we allow maximal occupancy $\nu\in \mathbb{N}$. Then, the configuration space is

@@

    \Omega:=\left\{(n_{1},\ldots,n_{N})\in\mathbb{N}_0^{N}\;:\; \sum_{k=1}^{N}n_{k}=\twoj\right\}^{|V|}

The edge dynamics consists in an exchange of particles between connected vertices. Moreover, each vertex can be connected with a reservoir that exchanges particles with the external environement. Mathematically, this can be described by an infinitesimal generator of the form
\begin{equation}\label{Generator}
    \mathcal{L}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\mathcal{L}_{x}
\end{equation}
where the quantities $ \omega_{x,y}\geq 0$ are called conductances and $\Gamma_{x}\geq 0$ local inhomogeneities. The generator $\mathcal{L}_{x,y}$ is called the \textit{edge generator}, while $\mathcal{L}_{x}$ is called the \textit{site generator}. These linear operators act on functions $f:\Omega\to \mathbb{R}$ as follows
\begin{equation}
\mathcal{L}_{x,y}f(n)=\sum_{k,l=1}^{N}n_{k}^{x}n_{l}^{y}\left[f(n-\delta_{x}^{k}+\delta_{l}^{x}+\delta_{k}^{y}-\delta_{l}^{y})-f(n)\right]
\end{equation}
\begin{equation}
    \mathcal{L}_{x}f(n)=\sum_{k,l=1}^{N}\alpha_{k}^{x}n_{l}^{x}\left[f(n+\delta_{k}^{x}-\delta_{l}^{x})-f(n)\right]
\end{equation}
where 
$\delta_{k}^{x}=(0,\ldots,0,1,0,\ldots,0)$ withe the one in the $k$-th position.
This means that on an edge at rate $\omega_{x,y}n_{k}^{x}n_{l}^{y}$ a particle of type $k$ at site $x$ is exchanged with a particle of tipe $l$ at site $y$; furthermore, on each site $x\in V$ at rate $\Gamma_{x}\alpha_{k}^{x}n_{l}^{x}$ a particle of type $l$ is replaced with a particle of type $k$. 
The following Figure \ref{fig:1}, Figure \ref{fig:2} help in understanding the dynamics. 
\begin{figure}
    \centering
    \includegraphics[scale=0.6]{Dyn_stir.eps}
    \caption{The edge dynamics}
    \label{fig:1}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[scale=0.88]{Dyn_stir_bordo.eps}
    \caption{The site dynamics}
    \label{fig:2}
\end{figure}\newline \\
{\color{blue}\textbf{Remark}: 
    We can always assign the name "empty" to one of the species of particles in the above dynamics (let's say, without loss of generality, the species 1). This species would play the role of "compensator" in the exclusion constraint, in the sense that its occupation will always be determined once we know the others, i.e. $\forall x\in V$: 
    \begin{equation}
        n_{1}^{x}=\twoj-\sum_{k=2}^{N}n_{k}^{x}
    \end{equation}
    Moreover, if we set $N=2$ and $\twoj=2j$, we exactly retrieve the SEP(2j), where the empty occupation is indeed $2j-\eta_{x}$ where $\eta_{x}$ is the number of (color blind) particles in the site $x$. In this sense we can think about the stirring process as the natural generalization of the SEP(2j).}
    \begin{flushright}
        $\square$
    \end{flushright}
The process described by the generator \eqref{Generator} is reversible if and only if, for any $k\in\{1,\ldots,N\}$, the boundary parameters $\alpha_{k}^{x}=\alpha_{k}$ $\forall x\in V$. The non-homogeneus reversible product measure has marginals the
\begin{equation}\label{reversibleMeasure}
    \mu^{rev}=\bigotimes_{x\in V}\mu^{rev}_{x}\qquad \mu_{x}^{rev}\sim \text{Multinomial}\left(\twoj,p_{1},\ldots,p_{N}\right)\quad \text{where}\quad p_{k}=\frac{\alpha_{k}}{\sum_{i=1}^{N}\alpha_{i}}
\end{equation}
This can be proven just by imposing the detailed balance conditions for the edge and for the site generators. Let us notice that the reversible measure is product only over the sites, but not over the species.
\subsection{The Lie algebraic description}

Consider the $gl(N)$ Lie algebra with with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{bc}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}
\end{equation}
{\color{red}The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu$. We assume $\lambda_i\geq \lambda_{i+1}$ without loss of generality.}
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \qquad\text{where}\qquad \twoj\in\mathbb{N}
\end{equation} 
The dimension $M$ of the symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition:
\begin{equation}
	M= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}
\end{equation} 
The basis elements of the vector space of the representation $\mathbb{C}^{M}$ are the column vectors 
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle
\end{equation}
%where $\sum_{k=1}^{N}n_{i}=\nu$. 
This basis vector satisfy the orthogonality relation
\begin{equation}\label{ortho}
    \langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{k=1}^{N}\delta_{m_{k},n_{k}}
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vector obtained by transposing $|m\rangle$ and $\delta_{m_{k},n_{k}}$ is the Kronhecker delta. 
The explicit actions of the algebra generators on the vectors $|n\rangle$ are the following:
\begin{equation}\label{actionE}
	\begin{cases}
		E_{ab}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle =n_{b}|n_{1},\ldots,n_{a}+1,\ldots,n_{b}-1,\ldots,n_{N}\rangle\quad a\neq b\\[0.1cm]
		E_{aa}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle = n_{a} |n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle\quad a=b
	\end{cases}
\end{equation}  
One can check that the generators defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield Dynkin weight \eqref{eq:dynkin}. The Cartan element $E_{aa}$ conserves the number of particles of type $a$.

\noindent
\textbf{Remark}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
\begin{equation}
   \bar E_{ab}=\mu\delta_{ab}-E_{N-b+1,N-a+1}
\end{equation}
{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
By the Lie algebra above we describe the process characterized by the generator \eqref{Generator}. We consider the state space \eqref{stateSpace} composed by

%We now define the equivalent of \eqref{stateSpace} in the vector notation
% \begin{equation}
%	\Omega':=\left\{|n\rangle=|n_{1},\ldots,n_{N}\rangle \;:\;n\in\mathbb{N}_0^N,\;\;|n|=\twoj\right\}^{\otimes|V|}
%	\end{equation}
%where we denote 
\begin{equation}
|{\bf n}\rangle=\left(\,\bigotimes_{x\in V}	|n_{1}^{x},\ldots,n_{N}^{x}\rangle\right)
\end{equation}
such that, $\forall x\in V$ we have $\sum_{k=1}^{N}n_{k}^{x}=\nu$. Moreover, the following orthogonality relations holds:
\begin{equation}
    \langle {\bf n}|{\bf m}\rangle =\prod_{x=1}^L\prod_{i=1}^N\delta_{n^x_i,m^{x}_i}
\end{equation}
We introduce the Hamiltonian operator

We can write this Hamiltonian in function of the coproduct of the second Casimir
\begin{equation}
    C_{2}=\sum_{a,b=1}^{N}E_{ab}E_{ba}
\end{equation}
It reads
\begin{equation}
	H_{x,y}=\left\{\frac{1}{2}\Delta^{xy}(C_{2})-\twoj(2\twoj+N)\frac{1}{2}\mathbbm{1}^{x}\otimes\mathbbm{1}^{y}\right\}
\end{equation}
Here we introduced the standard coproduct
\begin{equation}
\begin{split}
\Delta:gl(N)&\to gl(N)\otimes gl(N)\\
E_{ab}&\to E_{ab}\otimes \mathbbm{1}+\mathbbm{1}\otimes E_{ab}
\end{split}
\end{equation} acting on sites $x,y$ and used that $C_{2}|n\rangle=\twoj(\twoj+N)|n\rangle$.

Following  \cite{belitsky2015self}, the Hamiltonian and the genertor are linked by the following relation:
\begin{equation}
H=\mathcal{L}^{T}
\end{equation}
The generator can also be expressed as

\begin{equation}
    \mathcal{L}f( {\bf n})=\langle f|H| {\bf n}\rangle
\end{equation}
where 
\begin{equation}
    \langle f|=\sum_{ {\bf n}}f( {\bf n})\langle  {\bf n}|
\end{equation}

>>>>

At each site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $k$-th component $n_{k}^{x}$ denotes the number of particles of species $k\in \{1,\ldots,N\}$ at site $x\in V$. On each vertex we allow a total  maximal occupancy $\nu\in \mathbb{N}$. Then, the configuration space of the process on the graph $G$ is 

@@

    \Omega:=\left\{(n_{1},\ldots,n_{N})\in\mathbb{N}_0^{N}\;:\; \sum_{k=1}^{N}n_{k}=\twoj\right\}^{V}
Moreover, we denote a particle configuration as $\mathbf{n}\in \Omega$, where $\mathbf{n}=(n_{k}^{x})_{x\in V,\,k\in\{1,\ldots,N\}}$ .\\
The dynamics consists in an exchange of particles (at Poissonian times) between connected vertices. Moreover, each vertex  exchanges particles with the external environement. The infinitesimal generator of the process reads
\begin{equation}\label{Generator}
    \mathcal{L}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\mathcal{L}_{x}
where the quantities $ \omega_{x,y}\geq 0$ are called conductances and $\Gamma_{x}\geq 0$ local inhomogeneities. The generator $\mathcal{L}_{x,y}$ is called the \textit{edge generator}, while $\mathcal{L}_{x}$ is called the \textit{site generator}. These linear operators act on functions $f:\Omega\to \mathbb{R}$ as follows
\begin{equation}\label{edgeGenerator}
\mathcal{L}_{x,y}f(n)=\sum_{k,l=1}^{N}n_{k}^{x}n_{l}^{y}\left[f(n-\delta_{x}^{k}+\delta_{l}^{x}+\delta_{k}^{y}-\delta_{l}^{y})-f(n)\right]
\begin{equation}\label{siteGenerator}
    \mathcal{L}_{x}f(n)=\sum_{k,l=1}^{N}\alpha_{k}^{x}n_{l}^{x}\left[f(n+\delta_{k}^{x}-\delta_{l}^{x})-f(n)\right]
(\delta_{k}^{x})^{y}_{\ell}=\begin{cases}
1\qquad &\text{if}\quad y=x,\;\ell=k\\
0\qquad &\text{otherwise}
This means that on the edge $(x,y)\in \mathcal{E}$ at rate $\omega_{x,y}n_{k}^{x}n_{l}^{y}$ a particle of type $k$ at site $x$ is exchanged with a particle of tipe $l$ at site $y$; furthermore, on each site $x\in V$ at rate $\Gamma_{x}\alpha_{k}^{x}n_{\ell}^{x}$ a particle of type $\ell$ is replaced with a particle of type $k$. 
See Figure \ref{fig:1}, Figure \ref{fig:2}for a visual description of the dyamics. 
\begin{figure}
    \centering
    \includegraphics[scale=0.6]{Dyn_stir.eps}
    \caption{The edge dynamics}
    \label{fig:1}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[scale=0.88]{Dyn_stir_bordo.eps}
    \caption{The site dynamics}
    \label{fig:2}
\end{figure}\newline \\
{\color{blue}\textbf{Remark}: 
    We can always assign the name "empty" to one of the species of particles in the above dynamics (let's say, without loss of generality, the species 1). This species would play the role of "compensator" in the exclusion constraint, in the sense that its occupation will always be determined once we know the others, i.e. $\forall x\in V$: 
        n_{1}^{x}=\twoj-\sum_{k=2}^{N}n_{k}^{x}
    Moreover, if we set $N=2$ and $\twoj=2j$, we exactly retrieve the SEP(2j), where the empty occupation is indeed $2j-\eta_{x}$ where $\eta_{x}$ is the number of (color blind) particles in the site $x$. In this sense we can think about the stirring process as the natural generalization of the SEP(2j).}
The process described by the generator \eqref{Generator} is reversiblewith respect to the homogeneus product measure \begin{equation}\label{reversibleMeasure}
\mu_{rev}=\bigotimes_{x\in V}\mu^{rev}_{x}
when for any $k\in\{1,\ldots,N\}$
\begin{equation}\label{reversibilityCondition}
\alpha_{k}^{x}=\alpha_{k}\qquad \forall x\in V
This measure has marginals $\mu_{rev}^{x}$ distributed as 
 \mu^{x}_{rev}\sim \text{Multinomial}\left(\twoj,p_{1},\ldots,p_{N}\right)\quad \text{where}\quad p_{k}=\frac{\alpha_{k}}{\sum_{i=1}^{N}\alpha_{i}}
namely
\mu_{rev}^{x}(n^{x})=\frac{\nu}{\prod_{k=1}^{N}n_{k}^{x}}\prod_{k=1}^{N}p_{k}^{n_{k}^{x}}
This can be proven just by imposing the detailed balance conditions for the edge and for the site generators. If condition \eqref{reversibilityCondition} is not met, then reversibility is lost: each reservoir at site $x$ wants to fix its own density $\alpha^{x}=(\alpha_{1}^{x},\ldots,\alpha_{N}^{x})$ and currents will arise when at least two of them are different. 
\subsection{The Lie algebraic description}

Consider the $gl(N)$ Lie algebra with with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{bc}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu$. We assume $\lambda_i\geq \lambda_{i+1}$ without loss of generality.
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \qquad\text{where}\qquad \twoj\in\mathbb{N}
\end{equation} 
The dimension $M$ of the symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}
\end{equation} 
The basis elements of the vector space $\mathbb{C}^{M}$ are the column vectors 
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{sucht that}\quad \sum_{i=1}^{N}n_{i}=\nu
\end{equation}
%where $\sum_{k=1}^{N}n_{i}=\nu$. 
This basis vector satisfy the orthogonality relation with respect to the Euclidean scalar product 
\begin{equation}\label{ortho}
    \langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{k=1}^{N}\delta_{m_{k},n_{k}}
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vector obtained by transposing $|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{k},n_{k}}$ is the Kronhecker delta. 
The explicit actions of the algebra generators on the vectors $|n\rangle$ are the following:
\begin{equation}\label{actionE}
	\begin{cases}
		E_{ab}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle =n_{b}|n_{1},\ldots,n_{a}+1,\ldots,n_{b}-1,\ldots,n_{N}\rangle\quad a\neq b\\[0.1cm]
		E_{aa}|n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle = n_{a} |n_{1},\ldots,n_{a},\ldots,n_{b},\ldots,n_{N}\rangle\quad a=b
	\end{cases}
\end{equation}  
One can check that the matrices defined in this way satisfy the commutation relations \eqref{eq:comgl} and yield Dynkin weight \eqref{eq:dynkin}.

\noindent
\textbf{Remark}: The symmetric representations $\lambda=(\twoj,0,\ldots,0)$ are dual to the representations $\lambda=(\twoj,\ldots,\twoj,0)$ that can be obtained via 
\begin{equation}
   \bar E_{ab}=\mu\delta_{ab}-E_{N-b+1,N-a+1}
\end{equation}
{\color{red} This needs to be checked and give alternative form of Hamiltonian?}\\
By the Lie algebra above we describe the process characterized by the generator \eqref{Generator}. The state space \eqref{stateSpace} is the vector space with basis elements

%We now define the equivalent of \eqref{stateSpace} in the vector notation
% \begin{equation}
%	\Omega':=\left\{|n\rangle=|n_{1},\ldots,n_{N}\rangle \;:\;n\in\mathbb{N}_0^N,\;\;|n|=\twoj\right\}^{\otimes|V|}
%	\end{equation}
%where we denote 
\begin{equation}
|{\bf n}\rangle=\left(\,\bigotimes_{x\in V}	|n_{1}^{x},\ldots,n_{N}^{x}\rangle\right)
\end{equation}
where we require that $n_{k}^{x}\in \mathbb{N}_{0}$ and $\forall x\in V$ we have $\sum_{k=1}^{N}n_{k}^{x}=\nu$. The following orthogonality relation is a consequence of the single site relation \eqref{ortho}
\begin{equation}
    \langle {\bf n}|{\bf m}\rangle =\prod_{x=1}^L\prod_{i=1}^N\delta_{n^x_i,m^{x}_i}
\end{equation}
We introduce the Hamiltonian operator

\begin{equation}\label{OriginalHamiltonian}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}\label{edgeHamiltonian}
H_{x,y}=\sum_{k,l=1}^{N}\left(E_{kl}^{x}\otimes E_{lk}^{y}-E_{ll}^{x}\otimes E_{kk}^{y}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}\label{siteHamiltonian}
H_{x}=\sum_{k,l=1}^{N}\alpha_{k}^{x}\left(E_{kl}^{x}-E_{ll}^{x}\right)
\end{equation}
Here $E_{kl}^{x}$ is a copy of $E_{kl}$ defined in \eqref{actionE} acting on site $x$. 
Following  \cite{belitsky2015self}, the Hamiltonian and the genertor are linked by the following relation:
\begin{equation}
H=\mathcal{L}^{T}
\end{equation}
The generator can also be expressed as

\begin{equation}
    \mathcal{L}f( {\bf n})=\langle f|H| {\bf n}\rangle
\end{equation}
where 
\begin{equation}
    \langle f|=\sum_{ {\bf n\in \Omega}}f( {\bf n})\langle  {\bf n}|
\end{equation}

We can write this Hamiltonian as a function of the coproduct of the second Casimir
\begin{equation}
    C_{2}=\sum_{a,b=1}^{N}E_{ab}E_{ba}
\end{equation}
that acts diagonally as $C_{2}|\mathbf{n}\rangle=\twoj(\twoj+N)|\mathbf{n}\rangle$. \\
 Considering two sites $x,y\in \mathcal{E}$ the standard coproduct operator is defined as 
\begin{equation}
\begin{split}
\Delta^{x,y}:gl(N)&\to gl(N)\otimes gl(N)\\
E_{ab}&\to E_{ab}^{x}\otimes \mathbbm{1}^{y}+\mathbbm{1}^{x}\otimes E_{ab}^{y}
\end{split}
\end{equation}
therefore, we have 
\begin{equation}
\Delta^{x,y}(C_{2})=\sum_{a,b=1}^{N}\Delta(E_{ab})\Delta(E_{ba})=2\sum_{a,b=1}^{N}E_{ab}^{x}\otimes E_{ba}^{y}+C_{2}^{x}\otimes \mathbbm{1}^{y}+\mathbbm{1}^{x}\otimes C_{2}^{y}
\end{equation}

One can check that 
\begin{equation}\label{hamiltonianCasimir}
	H_{x,y}=\frac{1}{2}\Delta^{xy}(C_{2})-\twoj(2\twoj+N)\mathbbm{1}^{x}\otimes\mathbbm{1}^{y}
\end{equation}
